The hydra and the rooster

### 2020-08-13 latex/chapter-large-sets.tex
proof simplification in L_alpha

\subsubsection{The  ordinal zero}

The base case is directly a consequence of the specification.

@@

\subsubsection{Successor ordinals}

Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
$\beta$ at $k+2$ to $0$. 
By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.

@@

\subsubsection{Limit ordinals}

@@

Lemma L_succ_eqn alpha i :  L_ (Succ alpha) i = L_  alpha (S i).
\end{Coqsrc}

Using these three lemmas as rewrite rules, we can prove more properties of the \texttt{L\_} family of functions, including its correctness w.r.t. the specification
\texttt{L\_spec}.

\begin{todo}
The proof of \texttt{L\_alpha.L\_ok\_lim} seems to be redundant with
\texttt{Large\_Sets.L\_lim\_ok}.  
\end{todo}

@@

Lemma L_omega : forall k, L_ omega%e0 k = S (2 * k)%nat.

@@

@@

Theorem \texttt{battle\_length\_std } of
Module~\href{../src/html/hydras.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L\_alpha$.

@@

By induction on $i$, we prove a simple expression of \texttt{H\_ (OF i)}, where 
\texttt{OF $i$}  is the $i$-th finite ordinal.

\begin{Coqsrc}
Lemma H_Fin : forall i k: nat,  H_ i k = (i+k)%nat.
Proof with eauto with E0.
  induction i.
  - intros; simpl OF; simpl; autorewrite with H_rw E0_rw ... 
  - intros ;simpl; autorewrite with H_rw E0_rw ... 
    rewrite IHi; lia. 
Qed.
\end{Coqsrc}

@@

Our next step would be to prove an exact formula for $H_{\omega^\omega}(k)$.
Since the canonical sequence of $\omega^\omega$ is composed of all the
$\omega^i$, we first need to express $H_{\omega^i}$ for every natural number $i$.

@@

t seems hard to capture the complexity of this function by looking only at this
``exact'' formula. 
Let us consider a simple example, the number $H_{\omega^3}(3)$.

@@

On the contrary, the fonctions of the Hardy hierarchy have the following five properties. %%% ICI lien avec KS81

For any $\alpha < \epsilon_0$,
\begin{itemize}
\item the function $H_\alpha$ is strictly monotonous :
      For all $n,p \in\mathbb{N}, n < p \Rightarrow H_\alpha(n)< H_\alpha(p)$.
\item If $\alpha \not= 0$, then for every $n$, $n<H_\alpha(n)$.
\item The function $H_\alpha$ is pointwise less or equal than $H_{\alpha+1}$

>>>>

\subsubsection{The  ordinal zero}
\label{sect:L-spec-zero}
The base case is directly a consequence of the specification.

@@

\subsubsection{Successor ordinals}
\label{sect:L-spec-succ}
Let $\beta$ be some ordinal, and assume the arithmetic function $f$ satisfies 
the specification $(\texttt{L\_spec}\;\beta)$.  Let $k$ be any natural number.
Any path from $\texttt{succ}\,\beta$ to $0$ starting at $k+1$ can be decomposed into a first step from $\texttt{succ}\,\beta$ to $\beta$, then a path from
$\beta$ at $k+2$ to $0$. 
By hypothesis the interval $[k+2, f(k+2)-1]$ is $\beta$-mlarge.
But the interval $[k+1, f(k+2)-1]$ is the concatenation of the singleton
$\{k+1\}$ and the interval $[k+2, f(k+2)-1]$.
So, the function $\lambda\,k.\,f(k+1)$ satisfies the specification $\texttt{L\_spec}\,\beta$.

@@

\subsubsection{Limit ordinals}
\label{sect:L-spec-lim}

@@

Lemma L_succ_eqn alpha i :  L_ (Succ alpha) i = L_  alpha (S i).

Hint Rewrite L_zero_eqn L_succ_eqn : L_rw.
\end{Coqsrc}

Using these three lemmas as rewrite rules, we can prove more properties of the functions \texttt{L\_ $\alpha$}.

\begin{Coqsrc}
Lemma L_finite : forall i k :nat,  L_ i k = (i+k)%nat.  
(* Proof by induction on i, using L_zero_eqn and L_succ_eqn *)

Lemma L_omega : forall k, L_ omega%e0 k = S (2 * k)%nat.
(* Proof using L_finite and L_lim_eqn *)
\end{Coqsrc}

By  well-founded induction on $\alpha$, we prove the following lemmas:

@@

@@

Please note that the proof of \texttt{L\_correct} applies the lemmas proven in Sections~\ref{sect:L-spec-zero}, ~\ref{sect:L-spec-succ} and ~\ref{sect:L-spec-lim}.
Our previous study of \texttt{L\_spec} allowed us to pave the way for the definition by \texttt{Equations} and the correctness proof.

@@

Theorem \texttt{battle\_length\_std } of
Module~\href{../src/html/hydras.Hydra.Hydra_Theorems.html\#battle_length_std}{Hydra.Hydra\_Theorems} relates the length of standard battles with the functions $L_\alpha$.

@@

By induction on $i$, we prove a simple expression of \texttt{H\_ (Fin i)}, where 
\texttt{Fin $i$}  is the $i$-th finite ordinal.

\begin{Coqsrc}
Lemma H_Fin : forall i k: nat,  H_ (Fin i) k = (i+k)%nat.
Proof with eauto with E0.
  induction i.
  - intros; simpl OF; simpl; autorewrite with H_rw E0_rw ... 
  - intros ;simpl; autorewrite with H_rw E0_rw ... 
    rewrite IHi; lia. 
Qed.
\end{Coqsrc}

@@

Our next step would be to prove an exact formula for $H_{\omega^\omega}(k)$.
Since the canonical sequence of $\omega^\omega$ is composed of all the
$\omega^i$, we first need to express $H_{\omega^i}$ for any natural number $i$.

@@

\subsubsection{A numerical example}

It seems hard to capture the complexity of this function by looking only at this
``exact'' formula. 
Let us consider a simple example, the number $H_{\omega^3}(3)$.

@@

On the contrary, the fonctions of the Hardy hierarchy have the following five properties~\cite{KS81}: for any $\alpha < \epsilon_0$,
\begin{itemize}
\item the function $H_\alpha$ is strictly monotonous :
      For all $n,p \in\mathbb{N}, n < p \Rightarrow H_\alpha(n)< H_\alpha(p)$.
\item If $\alpha \not= 0$, then for every $n$, $n<H_\alpha(n)$.
\item The function $H_\alpha$ is pointwise less or equal than $H_{\alpha+1}$
